# Cloning a repository

## Prerequisites
- [Have an understanding of why Github is useful](https://hutchdatascience.org/dasl-snack-github)
- [Have a GitHub account](https://hutchdatascience.org/dasl-snack-github)
- [Create a repository on Github](https://hutchdatascience.org/dasl-snack-github)
- [Set up a Git Client](https://hutchdatascience.org/dasl-snack-github)

## What is cloning on GitHub?

In our daily grind, its ideal to work on your project from your own computer. To set this up, we'll need to `clone` it to our own computer. Cloning is making a remote copy of the project local.

```{r, fig.alt="To clone a repository means to copy a remote repository to your local computer", out.width = "100%", echo = FALSE}
ottrpal::include_slide("https://docs.google.com/presentation/d/1IJ_uFxJud7OdIAr6p8ZOzvYs-SGDqa7g4cUHtUld03I/edit#slide=id.g1014c75158f_0_90")
```

<div class = "github">
**clone** = To make a remote repository local. In other words, to make an online repository downloaded and linked on your computer.
</div>

To get started, you will need to clone the GitHub repository you created. We will be using this repository for the duration of this course.

```{r, fig.alt="To clone a GitHub repository, using GitKraken. First, Click Clone a repo. Then, choose where you’d like the repository to be on your computer using the ‘Browse’ button. Then Copy + Paste the url to the project you want to clone where it says ‘URL’. Then click `Clone the repository’.", out.width = "100%", echo = FALSE}
ottrpal::include_slide("https://docs.google.com/presentation/d/1IJ_uFxJud7OdIAr6p8ZOzvYs-SGDqa7g4cUHtUld03I/edit#slide=id.g1014c75158f_0_166")
```

It is simple to clone a GitHub repository using GitKraken. First, sign in to GitKraken; under Repository Management > Clone tab, click `Clone a repo`. Then, choose where you’d like the repository to be on your computer using the `Browse` button. You will need to `Copy + Paste` your new repository's URL (web address) to  where it says `URL`.

Navigate to your repository on GitHub to copy the URL. Copying and pasting is advisable because any little typo will inhibit cloning.

Now you are ready to click `Clone the repository`! It will ask you if you'd like to `Open Now`, click that.
